---
title: "HW7"
author: "Tianqi Wu"
date: "4/26/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r message=FALSE, warning=FALSE}
library(faraway)
library(leaps)
library(splines)
attach(prostate)
```


# Problem 1

## 1(a): Backward Elimination

If we set the p-value threshold to 15%, backward elimination selects the following predictors:  lcavol, lweight, lbph, svi. The adjusted R-squared is 0.6208.

```{r}
model.1a.1 = lm(lpsa~., data=prostate)
summary(model.1a.1)

model.1a.2 = lm(lpsa~.-gleason, data=prostate)
summary(model.1a.2)

model.1a.3 = lm(lpsa~.-gleason-lcp, data=prostate)
summary(model.1a.3)

model.1a.4 = lm(lpsa~.-gleason-lcp-pgg45, data=prostate)
summary(model.1a.4)

model.1a.5 = lm(lpsa~.-gleason-lcp-pgg45-age, data=prostate)
summary(model.1a.5)

anova(model.1a.5, model.1a.1)

```


## 1(b): AIC
lcavol, lweight, age, lbph, svi

```{r}
b=regsubsets(lpsa~., data=prostate)
rs = summary(b)

n=dim(prostate)[1]; msize = 2:9;
par(mfrow=c(2,2))
plot(msize, rs$adjr2, xlab="No. of Parameters", ylab = "Adjusted Rsquare");
plot(msize, rs$cp, xlab="No. of Parameters", ylab = "Mallow's Cp");

Aic = n*log(rs$rss/n) + 2*msize;
Bic = n*log(rs$rss/n) + msize*log(n);
plot(msize, Aic, xlab="No. of Parameters", ylab = "AIC");
plot(msize, Bic, xlab="No. of Parameters", ylab = "BIC")

## AIC
rs$which[which.min(Aic),]
```

## 1(b): BIC
lcavol, lweight, svi

```{r}
rs$which[which.min(Bic),]
```

## 1(c): Adjusted R^2
lcavol, lweight, age, lbph, svi, lcp, pgg45 
```{r}
rs$which[which.max(rs$adjr2),]
```

## 1(d): Mallows Cp
lcavol, lweight, lbph, svi
```{r}
rs$which[which.min(rs$cp),]
```

# Problem 2

## 2(a)
```{r}
## prepare data
set.seed(130)
fun<-function(x) sin(2*pi*x^3)^3
x<-seq(0,1,by=0.01)
y<-fun(x) + 0.1*rnorm(101)

## Regression Splines
model.2a = lm(y~bs(x, df=16, intercept=TRUE))
summary(model.2a)

plot(x, y)
lines(spline(x, predict(model.2a)), col="red", lty=1)
```

## 2(b)

AIC is -450.451 and BIC is -408.6091.

```{r}
rs = anova(model.2a)$`Sum Sq`[2]

n=101; msize = 16;
Aic = n*log(rs/n) + 2*msize;
Bic = n*log(rs/n) + msize*log(n)
Aic
Bic
```

## 2(c)

For Aic, the best model has 20 knots. For Bic, the best model has 15 knots.

```{r}
n=101; 
Aic = c(); Bic = c(); nparameter = c()
for (i in 3:20) {
  msize = i+4
  model.2c = lm(y~bs(x, df=msize, intercept = T))
  nparameter = append(nparameter, msize)
  rs = anova(model.2c)$`Sum Sq`[2]
  Aic = append(Aic, n*log(rs/n) + 2*msize)
  Bic = append(Bic, n*log(rs/n) + msize*log(n))
} 

plot(nparameter, Aic, xlab="No. of Parameters", ylab = "AIC");
plot(nparameter, Bic, xlab="No. of Parameters", ylab = "BIC")

nparameter[which.min(Aic)]
nparameter[which.min(Bic)]
```

## 2(d)

```{r}
## Aic
model.aic = lm(y~bs(x, df=20, intercept = T))

## Bic
model.bic = lm(y~bs(x, df=15, intercept = T))

plot(x, y)
lines(spline(x, predict(model.aic)), col="red", lty=1)
lines(spline(x, predict(model.bic)), col="blue", lty=1)
legend("topright", lty=rep(1,1), col=c("red", "blue"), legend=c("Aic", "Bic"))
```






